---
title: "REU Project"
format: html
editor: visual
---

### Set Up

```{r}
# Read in packages
library(dplyr)
library(stats)
library(multcompView)
library(ggpubr)
library(broom)
library(purrr)
```


## Data Wrangling
```{r}
# See GrowthAnalysis.R
```

## Analysis
```{r}
#alaxensis
salala_anova_model <- aov(log(total_dry_mass/total_fresh_length) ~ site, data = filter(results, Year.of.Growth == 2025, species == "salix_alaxensis"))
salala_tuk <- TukeyHSD(salala_anova_model)
salala_letters <- multcompLetters4(salala_anova_model, salala_tuk)
salala_cld_df <- data.frame(site = names(salala_letters$site$Letters),
                     Letters = salala_letters$site$Letters)
salala_cld_df

#pulchra
salpul_anova_model <- aov(log(total_dry_mass/total_fresh_length) ~ site, data = filter(results, Year.of.Growth == 2025, species == "salix_pulchra"))
salpul_tuk <- TukeyHSD(salpul_anova_model)
salpul_letters <- multcompLetters4(salpul_anova_model, salpul_tuk)
salpul_cld_df <- data.frame(site = names(salpul_letters$site$Letters),
                     Letters = salpul_letters$site$Letters)
salpul_cld_df
```

### Prep tukey for plot
```{r}
letter_df <- bind_rows(

  salala_cld_df %>%
    mutate(species = "salix_alaxensis"),

  salpul_cld_df %>%
    mutate(species = "salix_pulchra")

) %>%
  left_join(

    results %>%
      filter(
        Year.of.Growth == 2025,
        !is.na(total_dry_mass),
        !is.na(total_fresh_length),
        total_fresh_length > 0
      ) %>%
      mutate(
        growth = log(total_dry_mass / total_fresh_length)
      ) %>%
      group_by(site, species, region) %>%
      summarise(
        y = max(growth, na.rm = TRUE) + 0.12,
        .groups = "drop"
      ),

    by = c("site", "species")

  )
```



## Final Graphs

### Set up aesthetics
```{r}
my_theme <- theme_minimal() +
  theme(
    axis.text = element_text(size = 20),
    axis.title = element_text(size = 20),
    legend.title = element_text(size = 20),
    legend.text = element_text(size = 20),
    strip.text = element_text(size = 20))

my_colors <- scale_color_manual(
  values = c(
    "salix_alaxensis" = "darkgreen",
    "salix_pulchra" = "lightgreen"),
  labels = c(
    "salix_alaxensis" = "Salix alaxensis",
    "salix_pulchra" = "Salix pulchra"),
  name = "Species")

my_fill <- scale_fill_manual(
  values = c(
    "salix_alaxensis" = "darkgreen",
    "salix_pulchra" = "lightgreen"
  ),
  labels = c(
    "salix_alaxensis" = "Salix alaxensis",
    "salix_pulchra" = "Salix pulchra"
  ),
  name = "Species"
)
```

### Generate Plots
```{r}
p1 <- results %>%
  filter(
    Year.of.Growth == 2025,
    !is.na(total_dry_mass),
    !is.na(total_fresh_length),
    total_fresh_length > 0
  ) %>%
  mutate(site = forcats::fct_drop(site)) %>%
  ggplot(aes(
    x = site,
    y = log(total_dry_mass / total_fresh_length),
    color = species
  )) +
  geom_boxplot(linewidth = 1.15) +
  facet_wrap(~region, scales = "free_x") +
  scale_x_discrete(drop = TRUE) +
  my_colors +
  my_theme +
  xlab("Site") +
  ylab("Log(dry mass / fresh length)")
p1

p2 <- results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(
    x = species,
    y = log(total_dry_mass / total_fresh_length),
    fill = species
  )) +
  geom_boxplot() +
  my_fill +
  my_theme +
  xlab("Species") +
  ylab("Log(dry mass / fresh length)")
p2


p3 <- results_weather %>%
  filter(
    !is.na(species),
    !is.na(total_dry_mass),
    !is.na(total_fresh_length),
    !is.na(av_min_temp)
  ) %>%
  ggplot(aes(
    x = av_min_temp,
    y = log(total_dry_mass / total_fresh_length),
    color = species
  )) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  my_colors +
  my_theme +
  xlab("Average Minimum Temperature (°C)") +
  ylab("Log(dry mass / fresh length)")
p3

p4 <- results_weather %>%
  filter(
    !is.na(species),
    !is.na(total_dry_mass),
    !is.na(total_fresh_length),
    !is.na(av_light)
  ) %>%
  ggplot(aes(
    x = av_light,
    y = log(total_dry_mass / total_fresh_length),
    color = species
  )) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  my_colors +
  my_theme +
  xlab("Average Light") +
  ylab("Log(dry mass / fresh length)")
p4

p5 <- results %>%
  filter(
    Year.of.Growth == 2025,
    !is.na(species),
    !is.na(total_dry_mass),
    !is.na(total_fresh_length),
    !is.na(latitude)
  ) %>%
  ggplot(aes(
    x = latitude,
    y = log(total_dry_mass / total_fresh_length),
    color = species
  )) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  my_colors +
  my_theme +
  xlab("Latitude") +
  ylab("Log(dry mass / fresh length)")
p5
```

### Saving Plots
```{r}
ggsave("growth by site, species, region.png",
       plot = p1,
       width = 15,
       height = 10,
       dpi = 300)

ggsave("growth by species.png",
       plot = p2,
       width = 15,
       height = 10,
       dpi = 300)

ggsave("growth by min temp.png",
       plot = p3,
       width = 15,
       height = 10,
       dpi = 300)

ggsave("growth by light.png",
       plot = p4,
       width = 15,
       height = 10,
       dpi = 300)

ggsave("growth by latitude.png",
       plot = p5,
       width = 15,
       height = 10,
       dpi = 300)
```


### Regression Table
```{r}
temp_table <- results_weather %>%
  mutate(growth = log(total_dry_mass / total_fresh_length)) %>%
  filter(
    !is.na(species),
    !is.na(growth),
    !is.na(av_min_temp)
  ) %>%
  group_by(species) %>%
  do({
    fit <- lm(growth ~ av_min_temp, data = .)

    tibble(
      Intercept = round(coef(fit)[1], 3),
      Slope = round(coef(fit)[2], 3),
      R2 = round(summary(fit)$r.squared, 3),
      P_value = signif(summary(fit)$coefficients[2, 4], 3),
      N = nrow(.)
    )
  }) %>%
  mutate(Plot = "Average Minimum Temperature")

light_table <- results_weather %>%
  mutate(
    growth = log(total_dry_mass / total_fresh_length),
    av_light_1000 = av_light / 1000
  ) %>%
  filter(
    !is.na(species),
    !is.na(growth),
    !is.na(av_light_1000)
  ) %>%
  group_by(species) %>%
  do({
    fit <- lm(growth ~ av_light_1000, data = .)

    tibble(
      Intercept = round(coef(fit)[1], 3),
      Slope = round(coef(fit)[2], 3),
      R2 = round(summary(fit)$r.squared, 3),
      P_value = signif(summary(fit)$coefficients[2, 4], 3),
      N = nrow(.)
    )
  }) %>%
  mutate(Plot = "Average Light (per 1000 units)")

lat_table <- results %>%
  filter(Year.of.Growth == 2025) %>%
  mutate(growth = log(total_dry_mass / total_fresh_length)) %>%
  filter(
    !is.na(species),
    !is.na(growth),
    !is.na(latitude)
  ) %>%
  group_by(species) %>%
  do({
    fit <- lm(growth ~ latitude, data = .)

    tibble(
      Intercept = round(coef(fit)[1], 3),
      Slope = round(coef(fit)[2], 3),
      R2 = round(summary(fit)$r.squared, 3),
      P_value = signif(summary(fit)$coefficients[2, 4], 3),
      N = nrow(.)
    )
  }) %>%
  mutate(Plot = "Latitude")

regression_table <- bind_rows(
  temp_table,
  light_table,
  lat_table
)

regression_table
```

## Extra Visualizations

### Prelim
```{r}
# canopy height
results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(x = height_m, color = species)) +
  geom_density() +
  theme_minimal()

# canopy area
results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(x = ((canopy_width_m + canopy_length_m)/2)^2, color = species)) +
  geom_density() +
  theme_minimal()

# thaw
results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(x = thaw_base, color = species)) +
  geom_density() +
  theme_minimal()

results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(x = thaw_canopy, color = species)) +
  geom_density() +
  theme_minimal()

results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(x = thaw_canopy, y = thaw_base)) +
  geom_point() +
  geom_abline(intercept = 0, slope = 1, color = "black", linetype = "dashed")
  
```


### Box plots
```{r}
# alaxensis
results %>%
filter(Year.of.Growth == 2025, species == "salix_alaxensis") %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), color = site)) +
  geom_boxplot() +
  facet_wrap(~ region) +
  theme_minimal() +
  ylab("Salix Alaxensis log ( mass / length ) 2025")

# pulchra
results %>%
filter(Year.of.Growth == 2025, species == "salix_pulchra") %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), color = site)) +
  geom_boxplot() +
  facet_wrap(~ region) +
  theme_minimal() +
  ylab("Salix Pulchra log ( mass / length ) 2025")

# both
results %>%
filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), color = site)) +
  geom_boxplot() +
  facet_wrap(~ region) +
  theme_minimal() +
  ylab("Broad Trends log ( mass / length ) 2025")

results %>%
filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = species)) +
  geom_boxplot() +
  theme_minimal() +
  ylab("Salix Alaxensis log ( mass / length ) 2025")

```

### Environmental Vars with 2026 Field Data
```{r}
results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = thaw_base, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ region)
 
results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = thaw_canopy, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ region) 

results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = thaw_canopy, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = moisture_canopy, color = species)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE)

results %>%
  filter(Year.of.Growth == 2025) %>%
  ggplot(aes(y = (total_dry_mass/total_fresh_length), x = latitude)) +
  geom_point() +
  geom_smooth(method = "lm") +
  facet_wrap(~ species) +
  theme_minimal()
```

### Environmental Vars with Enviro 2025 Data
```{r}
### without p/R
results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_temp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  xlab("Average Temperature") +
  ylab("log ( mass / length )")

results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_min_temp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  xlab("Average Minimun Temperature") +
  ylab("log ( mass / length )")

results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_max_temp)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  xlab("Average Maximum Temperature") +
  ylab("log ( mass / length )")

results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_light)) +
  geom_point() +
  geom_smooth(method = "lm") +
  theme_minimal() +
  xlab("Average Light") +
  ylab("log ( mass / length )")


### with p/R
results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_temp)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation() +
  stat_cor(
    method = "pearson",
    #label.x = 11,
    label.y = -2.8
  ) +
  theme_minimal() +
  xlab("Average Temperature") +
  ylab("log ( mass / length )")

results_weather %>%
  ggplot(aes(y = log(total_dry_mass/total_fresh_length), x = av_min_temp)) +
  geom_point(size = 2) +
  geom_smooth(method = "lm", se = TRUE) +
  stat_regline_equation(
  aes(label = after_stat(eq.label)),
  size = 7
) +
  stat_cor(
    method = "pearson",
    label.y = -2.8,
    size = 7
  ) +
  theme_minimal() +
  xlab("Average Minimum Temperature") +
  ylab("log ( mass / length )") +
  theme_minimal() +
  theme(axis.text = element_text(size = 20),
        axis.title = element_text(size = 20),
        legend.title = element_text(size = 20),
        legend.text = element_text(size = 20),
        strip.text = element_text(size = 20))
```

```{r}
p1_extra <- results %>%
  filter(
    Year.of.Growth == 2025,
    !is.na(total_dry_mass),
    !is.na(total_fresh_length),
    total_fresh_length > 0
  ) %>%
  mutate(
    site = forcats::fct_drop(site),
    growth = log(total_dry_mass / total_fresh_length)
  ) %>%
  ggplot(aes(
    x = site,
    y = growth,
    color = species
  )) +

  geom_boxplot(
    linewidth = 1.15,
    width = 0.7,
    position = position_dodge(width = 0.75),
    outlier.shape = NA
  ) +

  geom_jitter(
    position = position_jitterdodge(
      jitter.width = 0.15,
      dodge.width = 0.75
    ),
    alpha = 0.5,
    size = 1.8
  ) +

  geom_text(
    data = letter_df,
    aes(
      x = site,
      y = y,
      label = Letters,
      color = species
    ),
    position = position_dodge(width = 0.75),
    size = 6,
    fontface = "bold",
    show.legend = FALSE
  ) +

  facet_wrap(~region, scales = "free_x") +

  scale_x_discrete(drop = TRUE) +

  my_colors +
  my_theme +

  xlab("Site") +
  ylab("Log(dry mass / fresh length)")